Generality in food webs scales with species richness, not latitude

### 2015-04-09 manuscript/latitudepaper.tex
Updated first part of methods

[[Breaking things down by TL:

  Using proportions in main text following Riede et. al.

  With the subset of webs that actually have >0 BIT, there is a relationship between S, Gen, Vul, B, and T and Lat. Looks like this might be driven by outliers though... so gotta test that but might get to ignore things again.

]]

@@

@@

We compiled a list of 263 empirical food webs from
multiple sources (see Supplemental Information for web origins and selection
criteria). We recorded study site latitude from the original source where
possible or, where study sites were described but exact coordinates were not
given, obtained estimated coordinates using Google Earth \citep{GoogleEarth}.
We then divided the species in each web into basal resources (those species with
consumers but no prey), top predators (those species with prey but no predators),
and intermediate consumers (species with both predators and prey, including cannibalistic
species). For our future analyses, we considered only those webs with at least one basal
resource, one intermediate consumer, and one top predator. This reduced our dataset to
163 food webs.

% [[list in supplemental (347 GlobalWeb webs \citep{GlobalWeb}), supplemented 
% with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
% and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
% Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
% webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
% competitive interactions (8, 34, and 2, respectively), described inferred interactions in an extinct community (2), were ``generalized 
% schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
% described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
% exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
% seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
% interactions between them \citep{Dunne2013}.

As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
Many analyses of food-web structure reduce this variation by aggregating species with identical predator and prey
sets to form ``trophic species'' webs (e.g. ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
of the dataset. As the version of the dataset used did not qualitatively change the results, we present only
the results for the original webs in the main text. [[so what does this suggest? i find that in methods when making these statements it helps to go the step beyond and try and infer why trophic species/SR yielded similar results]] In each case, we counted the number of species ($S$) and 
links in each web and used these to calculate the mean number of links per species ($Z$), mean generality 
($G$), and mean vulnerability ($V$) of each web.

@@

To determine whether there were latitudinal gradients in food-web structure,
we first examined simple linear relationships between latitude and $S$, $Z$, $G$, and $V$. 
We fit models of the form

\begin{equation}
\label{Latfull}
S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} 
\end{equation}

where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
latitude (degrees north or south  regardless of direction), $E_{i}$ is the
ecosystem type of network $i$ (comprising terms for stream, marine, lake, and terrestrial networks with estuarine
networks being the intercept) and $\epsilon_{i}$ a residual error term. 
We then simplified these maximal models. To do this, we calculated the AIC
of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models identified
using R~\citep{R} function dredge from packages MuMIn~\citep{MuMIn}. Simplified models were obtained by
systematically removing all possible combinations of terms from the full model.
The best-fitting model was then determined to be the model with the fewest terms 
where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
model.

@@

\begin{equation}
\label{Power}
Z_{i}~\alpha S_{i}^{\beta}
\end{equation}

which is often re-expreseed in logarithmic form

\begin{equation}
\label{Loglog}
\log{Z_{i}} ~ \log{\alpha} + \beta\log{S_{i}}
\end{equation}

Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
Specifically,~\ref{Power} implies a normally-distributed, additive error and~\ref{Loglog} a lognormal,
multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
over another, we follow the recommendations in \citet{Xiao2011} and compared the two
model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
data was then used to test for potential effects of latitude.

@@

  For relationships of $Z$, $G$, $V$, $B$, $I$, and $T$ to $S$, equation~\ref{Loglog} had a
  lower AIC than the corresponding equation~\ref{Power}. This indicates that the
  data support an assumption of multiplicative log-normal error better than an
  assumption of additive normal error. That is, models where $\epsilon$ is
  modelled as an exponent to a multiplicative term (that is, an additive term on
  the logarithmic scale) provide a better description of the data than models
  where $\epsilon$ is modelled as an additive term on the arithmetic scale
  [[That's what this means... not sure if explaining it actually adds anything
  though]] .  We therefore used logarithmic-form models when assessing the
  effect of latitude on scaling relationships  with $S$.

  Contrary to the expected latitudinal gradient, the best-fit version of
  equation~\ref{Latfull} for $S$ included a signficant effect of latitude on $S$
  in stream webs only ($\alpha_{Latitude:Stream}$=-1.69, $p$=0.007).  This
  relationship was robust to the removal of three outliers (post-removal
  $\alpha_{Latitude:Stream}$=-1.17, $p$=0.038). The best-fit versions of
  equation~\ref{Latfull} for $Z$ and $V$ also included relatioships with
  latitude in stream webs ($\alpha_{Latitude:Stream}$=-0.179, $p$=0.007 and
  $\alpha_{Latitude:Stream}$=-0.179, $p$=0.007, respectively). However, after
  the removal of three outlying stream webs with very high $Z$ and $V$ (two of
  which also had unusually high $S$) these relationships were no longer
  significant ($\alpha_{Latitude:Stream}$=-0-0.046, $p$=0.373 and
  $\alpha_{Latitude:Stream}$=-0.046, $p$=0.373, respectively). The best-fit
  version of equation~\ref{Latfull} for $G$ did not include any effect of
  latitude in any ecosystem type. Marine food webs
  had lower $S$, $Z$, $G$, and $V$ than all other web types (post-removal
  $\alpha_{Marine}$=-29.5, $p$\textless0.001, $\alpha_{Marine}$=-1.46,
  $p$=0.035, $\alpha_{Marine}$=-1.74, $p$=0.0612, and $\alpha_{Marine}$=-1.46,
  $p$=0.035, respectively) while stream food webs had higher $G$ than webs from
  other ecosystem types (post-removal $\alpha_{Stream}$=3.51,
  $p$\textless0.001) [[Check streams in $V$, $Z$ for post-removal models]].

@@

$Z$, $G$, and $V$ all increased with increasing $S$ at all latitudes. The strength of
this increase varied both over latitude and across ecosystem types. For $Z$,
scaling was stronger towards the equator in estuarine, marine, stream, and
terrestrial ecosystems (). In lake ecosystems, however, scaling was stronger
toward the poles. $G$ and $V$ followed the same trends, except that scaling of $G$
with $S$ was stronger towards the poles as in lakes ().

@@

In lake food webs, species in tropical food webs also gained fewer predators,
and therefore fewer links overall. Conversely, for species in high-latitude
food webs numbers of predators, prey, and links per species all increased
rapidly with increasing species richness. This may reflect prey switching as
different food sources become available in different seasons~\citep{}.
Alternatively, the sharper increases in links per species, generality, and
vulnerability at higher latitudes may be driven by the relationships between
each property and the proportion of basal resources in these food webs.
[[Except each property increases more with S and decreases more with B...]]

Surprisingly species in high-latitude stream food webs, while gaining many
additional prey species per additional species in the food web, did not
receive any more predators or links than species in low-latitude streams. The
limited influence of latitude on the scaling relationships of links per
species and vulnerability may be because, for these properties, latitude had
opposite effects on the scaling relationships with proportions of basal
resources and top predators. When considering generality, however, only the
proportion of basal resources varied significantly over latitude. [[Same
problem as in lakes, the trends are opposite.]]

@@

\begin{figure}[!h]
\label{LS}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/LS_nonts.eps}
\caption{Relationships between mean links per species and species-richness, proportion of basal resources, proportion of intermediate consumers, or top predators (N=163 food webs). [[More details needed]]}
\end{figure}

\begin{figure}[!h]
\label{Gen}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/Gen_nonts.eps}
\caption{Relationships between mean generality (number of prey) and species-richness, proportion of basal resources, proportion of intermediate consumers, or top predators (N=163 food webs). [[More details needed]]}
\end{figure}

\begin{figure}[!h]
\label{Vul}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/Vul_nonts.eps}
\caption{Relationships between mean vulnerability (number of predators) and species-richness, proportion of basal resources, proportion of intermediate consumers, or top predators (N=163 food webs). [[More details needed]]}
\end{figure}

>>>>

@@

[[Still need to check that new results hold for TS versions of webs.]]

@@

  We compiled a list of 163 empirical food webs from
  multiple sources (see Supplemental Information for web origins and selection
  criteria). We recorded study site latitude from the original source where
  possible or, where study sites were described but exact coordinates were not
  given, obtained estimated coordinates using Google Earth \citep{GoogleEarth}.
  We then divided the species in each web into basal resources (those species with
  consumers but no prey), top predators (those species with prey but no predators),
  and intermediate consumers (species with both predators and prey, including cannibalistic
  species).

  % [[list in supplemental (347 GlobalWeb webs \citep{GlobalWeb}), supplemented 
  % with 31 food webs used to study the relationship between consumer and resource body sizes (SizeWebs \citep{Brose2006})
  % and seven food webs used to study the impact of parasites on network structure (ParWebs \citep{Dunne2013}).
  % Of the GlobalWeb webs, 122 were rejected because their original source could not be found or was unpublished (53), they were source or sink
  % webs rather than descriptions of an entire community (13), were focused on plant-pollinator, host-parasitoid, or 
  % competitive interactions (8, 34, and 2, respectively), described inferred interactions in an extinct community (2), were ``generalized 
  % schemes'' rather than being based on empirical observation (8), or because it was not clear which of a variety of 
  % described sites the published food web represented (2). (See Table S1 for a list of rejected webs and reasons for their
  % exclusion.) For three GlobalWeb webs which included parasites as a minor component of the community, as well as the 
  % seven ParWebs webs, we included a modified version of each web which included only free-living species and the 
  % interactions between them \citep{Dunne2013}. 
  % As we were considering the log of proportions of basal resources, top predators, and intermediate consumers, each web used in our analysis had to have at least one species in each trophic group.

  As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
  is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
  Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
  sets to form ``trophic species'' webs (e.g. ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}). As this study
  is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
  species with redundant sets of interactions. We therefore analysed both original and trophic-species versions
  of the dataset; in each case using the number of species ($S$) and 
  links ($L$) in each web to calculate the mean number of links per species ($Z$), mean generality 
  ($G$), and mean vulnerability ($V$) of the web. 
  The version of the dataset used did not qualitatively change the results, suggesting that
  the scaling relationships between species richness, other food-web properties, and latitude are very 
  similar whether or not redundant species are included. For simplicity, we present only
  the results for the original webs in the main text.

@@

  To determine whether there were latitudinal gradients in food-web structure,
  we first examined simple linear relationships between latitude and each of $S$, $Z$, $G$, and $V$. 
  We fit models of the form

  \begin{equation}
  \label{Latfull}
  S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} .,
  \end{equation}

  \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
  latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
  variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
  marine, lake, and terrestrial networks with estuarine
  networks providing the intercept) and $\epsilon_{i}$ a residual error term. 
  We next calculated the AIC
  of the maximal model, as well as the AIC's of a suite of candidate simplified models identified
  using the R~\citep{R} function dredge from package MuMIn~\citep{MuMIn}. 
  Simplified models were obtained by
  systematically removing all possible combinations of terms from the full model.
  The best-fitting model was then determined to be the model with the fewest terms 
  where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
  and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
  model.

@@

  \begin{equation}
  \label{Power}
  Z_{i} \~ \alpha S_{i}^{\beta}
  \end{equation}

  \noindent which is often re-expressed in logarithmic form

  \begin{equation}
  \label{Loglog}
  \log{Z_{i}} \~ \log{\alpha} + \beta\log{S_{i}}
  \end{equation}

  Although these relationships are very similar, they imply different error distributions~\citep{Xiao2011}.
  Specifically, equation~(\ref{Power}) implies a normally-distributed, additive error and equation~(\ref{Loglog}) a lognormal,
  multiplicative error. As we have no \emph{a priori} reason to believe that our dataset has one error distribution
  over another, we follow the recommendations in \citet{Xiao2011} and compared the two
  model formulations explicitly. The model with the error distribution most resembling that observed in the empirical
  data was then used to test for potential effects of latitude.

@@

  When considering the relationships between $S$ and $Z$, $G$, $V$, $B$, $I$, and $T$,
  equation~(\ref{Loglog}) had a
  lower AIC than the corresponding equation~(\ref{Power}). This indicates that the
  data support an assumption of multiplicative log-normal error better than an
  assumption of additive normal error. That is, models where $\epsilon$ is
  modelled as an exponent to a multiplicative term (that is, an additive term on
  the logarithmic scale) provide a better description of the data than models
  where $\epsilon$ is modelled as an additive term on the arithmetic scale.  
  We therefore used logarithmic-form models when assessing the
  effect of latitude on scaling relationships  with $S$.

  Contrary to the expected latitudinal gradient, the best-fit version of
  equation~(\ref{Latfull}) for $S$ included a signficant effect of latitude on $S$
  in stream webs only ($\alpha_{Latitude:Stream}$=-1.69, $p$=0.007).  This
  relationship was robust to the removal of three outliers (post-removal
  $\alpha_{Latitude:Stream}$=-1.17, $p$=0.038). The best-fit versions of
  equation~(\ref{Latfull}) for $Z$ and $V$ also included relationships with
  latitude in stream webs ($\alpha_{Latitude:Stream}$=-0.179, $p$=0.007 and
  $\alpha_{Latitude:Stream}$=-0.179, $p$=0.007, respectively). However, after
  the removal of three outlying stream webs with very high $Z$ and $V$ (two of
  which also had unusually high $S$) these relationships were no longer
  significant ($\alpha_{Latitude:Stream}$=-0-0.046, $p$=0.373 and
  $\alpha_{Latitude:Stream}$=-0.046, $p$=0.373, respectively). The best-fit
  version of equation~(\ref{Latfull}) for $G$ did not include any effect of
  latitude in any ecosystem type. Marine food webs
  had lower $S$, $Z$, $G$, and $V$ than all other web types (post-removal
  $\alpha_{Marine}$=-29.5, $p$\textless0.001, $\alpha_{Marine}$=-1.46,
  $p$=0.035, $\alpha_{Marine}$=-1.74, $p$=0.0612, and $\alpha_{Marine}$=-1.46,
  $p$=0.035, respectively) while stream food webs had higher $G$ than webs from
  other ecosystem types (post-removal $\alpha_{Stream}$=3.51,
  $p$\textless0.001).

@@

$Z$, $G$, and $V$ all increased with increasing $S$ at all latitudes. The strength of
this increase varied both over latitude and across ecosystem types. For $Z$,
scaling was stronger towards the equator except in lake ecosystems where scaling was stronger
towards the poles (Fig.~\ref{LS}). $G$ and $V$ followed the same trends, except that scaling of $G$
with $S$ was stronger towards the poles as in lakes (Figs.~\ref{Gen, Vul}).

@@

   $B$ increased towards the poles ($\alpha_{Latitude:Stream}$=0.012, $p$\textless0.001)
   $T$ decreased towards the poles in stream food webs ($\alpha_{Latitude}$=0.001, $p$=0.136;
  $\alpha_{Latitude:Stream}$=-0.006, $p$=0.030)

In lake food webs, species in tropical food webs also gained fewer predators,
and therefore fewer links overall. Conversely, for species in high-latitude
food webs numbers of predators, prey, and links per species all increased
rapidly with increasing species richness.
As there were proportionally fewer top predators in high-latitude lakes yet
generality, vulnerability, and links per species all increased more rapidly than
near the equator, it appears that high-latitude predators are more likely to be 
generalists. This may reflect prey switching as
different food sources become available in different seasons~\citep{}.

Surprisingly species in high-latitude stream food webs, while gaining many
additional prey species per additional species in the food web, did not
receive any more predators or links than species in low-latitude streams. 
This may be partly due to the increasing proportion of the food web made up of
basal resources at higher latitudes. If any additional species is more likely to
be a basal resource at high latitudes, it will have predators but no prey.
Combined with the potential for prey-switching between seasonal resources, this 
could account for the sharper increase in generality with increasing species richness
in high-latitude streams. Vulnerability and mean links per species, on the other hand,
increased at the same rate across latitude. [[Still working on why]]

@@

\begin{figure}[!h]
\label{S}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/B_latlines_nonts.eps}
\caption{}
\end{figure}

\begin{figure}[!h]
\label{B}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/B_latlines_nonts.eps}
\caption{}
\end{figure}

\begin{figure}[!h]
\label{I}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/I_latlines_nonts.eps}
\caption{}
\end{figure}

\begin{figure}[!h]
\label{T}
\includegraphics[width=.9\textwidth]{Figures/by_TL/scaling_with_S/proportions/T_latlines_nonts.eps}
\caption{}
\end{figure}
